### 2023-06-15 figs/markov_state.tex
major work

%% \documentclass[crop,tikz]{standalone}
%% \input{defs}
%% \begin{document}
\begin{tikzpicture}[scale=1]
  \usetikzlibrary{shapes,arrows,positioning}
  \tikzstyle{coordinate}=[inner sep=0pt,outer sep=0pt]
  \tikzstyle{state}=[shape=ellipse, color=black, draw, font=\large, fill=white, thick, minimum size=5em]
  \tikzstyle{trans}=[color=black, thick, >=stealth]
  \node[state] (s0) at (0,0) {$x=(S,E,I,R)$};
  \node[state] (s1) at (165:7.8) {$x'=(S-1,E+1,I,R)$};
  \node[state] (s2) at (195:7.1) {$x'=(S,E-1,I+1,R+1)$};
  \node[state] (s3) at (295:4) {$x'=(S,E,I-1,R+1)$};
  \node[state] (s4) at (355:9) {$x'=(S+1,E,I,R-1)$};
  \node[state] (s5) at (18:6) {$x'=(S,E,I,R)$};
  \draw[trans,->] (s0) -- (s1) node[midway,above,sloped] {transmission};
  \draw[trans,->] (s0) -- (s2) node[midway,above,sloped] {progression};
  \draw[trans,->] (s0) -- (s3) node[midway,above,sloped] {recovery};
  \draw[trans,->] (s0) -- (s4) node[midway,above,sloped] {loss of immunity};
  \draw[trans,->] (s0) -- (s5) node[midway,above,sloped] {sampling};
\end{tikzpicture}
%% \end{document}