### 2024-05-10 figs/event_types.tex
more

\begin{center}
  \resizebox{0.7\linewidth}{!}{
    \begin{tikzpicture}[scale=2]
      \usetikzlibrary{shapes,arrows,positioning}
      \usetikzlibrary{arrows.meta,positioning,decorations}
      \tikzstyle{label}=[font=\Large]
      \tikzstyle{coordinate}=[inner sep=0pt,outer sep=0pt]
      \tikzstyle{deme1}=[color=darkblue, line width=3pt]
      \tikzstyle{deme1H}=[color=darkblue, line width=3pt, densely dotted]
      \tikzstyle{deme2}=[color=maize, line width=3pt]
      \tikzstyle{deme2H}=[color=maize, line width=3pt, densely dotted]
      \tikzstyle{axis}=[color=black, thick, >=stealth]
      \tikzstyle{timepoint}=[color=red!70!black]

      \def\ysp{0.2} %% spacing between adjacent lineages

      %% Birth-type event
      \begin{scope}[shift={(0,0)}]
        \coordinate (origin) at (0,0);
        \node[label] (lab) at (origin) {\textbf{A}};
        \coordinate (left) at ($(origin)+(0.1,-0.1)$);
        \coordinate (right) at ($(left)+(1,0)$);
        \coordinate (mid) at ($(left)!0.5!(right)$);
        \coordinate (rlab) at ($(right)+(0.1,0)$);
        \coordinate (bleft) at ($(left)+(0,-8*\ysp)$);
        \coordinate (bright) at (rlab |- bleft);
        \coordinate (temp) at ($(origin)+(0,-4*\ysp)$);
        \coordinate (temp1) at ($(temp)+(0,\ysp)$);
        \coordinate (temp2) at ($(temp)-(0,\ysp)$);
        \draw[deme2] (mid |- temp) -- (mid |- temp1) -- (right |- temp1);
        \draw[deme1] (mid |- temp) -- (mid |- temp2) -- (right |- temp2);
        \draw[deme1] (left |- temp) -- (right |- temp);
        \node[text=darkgreen,font=\Huge] at (mid |- temp) {$\bullet$};
        \foreach \y in {1,7} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme1] (left |- temp) -- (right |- temp);
        }
        \foreach \y in {6,8} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme2] (left |- temp) -- (right |- temp);
        }
        \foreach \y in {2,7} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme2] (left |- temp) -- (right |- temp);
        }
        \draw[timepoint] (mid) -- (mid |- bleft);
        \draw[axis,->] (bleft) -- (bright);
        \node[xshift=1ex] (tlab) at (bright.east) {$t$};
      \end{scope}

      %% Death-type event
      \begin{scope}[shift={(2,0)}]
        \coordinate (origin) at (0,0);
        \node[label] (lab) at (origin) {\textbf{B}};
        \coordinate (left) at ($(origin)+(0.1,-0.1)$);
        \coordinate (right) at ($(left)+(1,0)$);
        \coordinate (mid) at ($(left)!0.5!(right)$);
        \coordinate (rlab) at ($(right)+(0.1,0)$);
        \coordinate (bleft) at ($(left)+(0,-8*\ysp)$);
        \coordinate (bright) at (rlab |- bleft);
        \foreach \y in {1,2,6} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme1] (left |- temp) -- (right |- temp);
        }
        \foreach \y in {3,8} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme2] (left |- temp) -- (right |- temp);
        }
        \coordinate (temp) at ($(origin)+(0,-4*\ysp)$);
        \draw[deme2H] (left |- temp) -- (mid |- temp);
        \node[text=red,font=\Huge] at (mid |- temp) {$\times$};
        \foreach \y in {5,7} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme2] (left |- temp) -- (right |- temp);
        }
        \draw[timepoint] (mid) -- (mid |- bleft);
        \draw[axis,->] (bleft) -- (bright);
        \node[xshift=1ex] (tlab) at (bright.east) {$t$};
      \end{scope}

      %% Migration-type event
      \begin{scope}[shift={(4,0)}]
        \coordinate (origin) at (0,0);
        \node[label] (lab) at (origin) {\textbf{C}};
        \coordinate (left) at ($(origin)+(0.1,-0.1)$);
        \coordinate (right) at ($(left)+(1,0)$);
        \coordinate (mid) at ($(left)!0.5!(right)$);
        \coordinate (rlab) at ($(right)+(0.1,0)$);
        \coordinate (bleft) at ($(left)+(0,-8*\ysp)$);
        \coordinate (bright) at (rlab |- bleft);
        \foreach \y in {2,3,8} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme1] (left |- temp) -- (right |- temp);
        }
        \coordinate (temp) at ($(origin)+(0,-4*\ysp)$);
        \draw[deme2] (left |- temp) -- (mid |- temp);
        \draw[deme1] (mid |- temp) -- (right |- temp);
        \node[text=darkgreen,font=\Huge] at (mid |- temp) {$\bullet$};
        \foreach \y in {1,5} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme2] (left |- temp) -- (right |- temp);
        }
        \foreach \y in {6,7} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme1] (left |- temp) -- (right |- temp);
        }
        \draw[timepoint] (mid) -- (mid |- bleft);
        \draw[axis,->] (bleft) -- (bright);
        \node[xshift=1ex] (tlab) at (bright.east) {$t$};
      \end{scope}

      %% Sample event
      \begin{scope}[shift={(6,0)}]
        \coordinate (origin) at (0,0);
        \node[label] (lab) at (origin) {\textbf{D}};
        \coordinate (left) at ($(origin)+(0.1,-0.1)$);
        \coordinate (right) at ($(left)+(1,0)$);
        \coordinate (mid) at ($(left)!0.5!(right)$);
        \coordinate (rlab) at ($(right)+(0.1,0)$);
        \coordinate (bleft) at ($(left)+(0,-8*\ysp)$);
        \coordinate (bright) at (rlab |- bleft);
        \foreach \y in {2,5} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme1] (left |- temp) -- (right |- temp);
        }
        \foreach \y in {1,4,7} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme2] (left |- temp) -- (right |- temp);
        }
        \foreach \y in {3,6,8} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme2] (left |- temp) -- (right |- temp);
        }
        \coordinate (temp) at ($(origin)+(0,-2*\ysp)$);
        \node[text=royalblue,font=\Huge] at (mid |- temp) {$\bullet$};
        \coordinate (temp) at ($(origin)+(0,-6*\ysp)$);
        \node[text=royalblue,font=\Huge] at (mid |- temp) {$\bullet$};
        \draw[timepoint] (mid) -- (mid |- bleft);
        \draw[axis,->] (bleft) -- (bright);
        \node[xshift=1ex] (tlab) at (bright.east) {$t$};
      \end{scope}

      %% Neutral-type event
      \begin{scope}[shift={(0,-2)}]
        \coordinate (origin) at (0,0);
        \node[label] (lab) at (origin) {\textbf{E}};
        \coordinate (left) at ($(origin)+(0.1,-0.1)$);
        \coordinate (right) at ($(left)+(1,0)$);
        \coordinate (mid) at ($(left)!0.5!(right)$);
        \coordinate (rlab) at ($(right)+(0.1,0)$);
        \coordinate (bleft) at ($(left)+(0,-8*\ysp)$);
        \coordinate (bright) at (rlab |- bleft);
        \foreach \y in {4,6,7} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme1] (left |- temp) -- (right |- temp);
        }
        \foreach \y in {1,2,5} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme2] (left |- temp) -- (right |- temp);
        }
        \foreach \y in {3,8} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme1] (left |- temp) -- (right |- temp);
        }
        \draw[timepoint] (mid) -- (mid |- bleft);
        \draw[axis,->] (bleft) -- (bright);
        \node[xshift=1ex] (tlab) at (bright.east) {$t$};
      \end{scope}

      %% Compound birth/death (Moran) event
      \begin{scope}[shift={(2,-2)}]
        \coordinate (origin) at (0,0);
        \node[label] (lab) at (origin) {\textbf{F}};
        \coordinate (left) at ($(origin)+(0.1,-0.1)$);
        \coordinate (right) at ($(left)+(1,0)$);
        \coordinate (mid) at ($(left)!0.5!(right)$);
        \coordinate (rlab) at ($(right)+(0.1,0)$);
        \coordinate (bleft) at ($(left)+(0,-8*\ysp)$);
        \coordinate (bright) at (rlab |- bleft);
        \coordinate (temp) at ($(origin)+(0,-4*\ysp)$);
        \coordinate (temp1) at ($(temp)+(0,\ysp)$);
        \draw[deme1] (mid |- temp) -- (mid |- temp1) -- (right |- temp1);
        \draw[deme1] (left |- temp) -- (right |- temp);
        \draw[deme2H] (left |- temp1) -- (mid |- temp1);
        \node[text=darkgreen,font=\Huge] at (mid |- temp) {$\bullet$};
        \node[text=red,font=\Huge] at (mid |- temp1) {$\times$};
        \foreach \y in {2,5} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme1] (left |- temp) -- (right |- temp);
        }
        \foreach \y in {1,7} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme2] (left |- temp) -- (right |- temp);
        }
        \foreach \y in {6,8} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme1] (left |- temp) -- (right |- temp);
        }
        \draw[timepoint] (mid) -- (mid |- bleft);
        \draw[axis,->] (bleft) -- (bright);
        \node[xshift=1ex] (tlab) at (bright.east) {$t$};
      \end{scope}

      %% Compound sample/death event
      \begin{scope}[shift={(4,-2)}]
        \coordinate (origin) at (0,0);
        \node[label] (lab) at (origin) {\textbf{G}};
        \coordinate (left) at ($(origin)+(0.1,-0.1)$);
        \coordinate (right) at ($(left)+(1,0)$);
        \coordinate (mid) at ($(left)!0.5!(right)$);
        \coordinate (rlab) at ($(right)+(0.1,0)$);
        \coordinate (bleft) at ($(left)+(0,-8*\ysp)$);
        \coordinate (bright) at (rlab |- bleft);
        \coordinate (temp) at ($(origin)+(0,-5*\ysp)$);
        \draw[deme1] (left |- temp) -- (mid |- temp);
        \node[text=royalblue,font=\Huge] at (mid |- temp) {$\bullet$};
        \node[text=red,font=\Huge] at (mid |- temp) {$\times$};
        \foreach \y in {1,6} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme1] (left |- temp) -- (right |- temp);
        }
        \foreach \y in {2,3,8} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme2] (left |- temp) -- (right |- temp);
        }
        \foreach \y in {4,7} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme2] (left |- temp) -- (right |- temp);
        }
        \draw[timepoint] (mid) -- (mid |- bleft);
        \draw[axis,->] (bleft) -- (bright);
        \node[xshift=1ex] (tlab) at (bright.east) {$t$};
      \end{scope}

      %% Compound event
      \begin{scope}[shift={(6,-2)}]
        \coordinate (origin) at (0,0);
        \node[label] (lab) at (origin) {\textbf{H}};
        \coordinate (left) at ($(origin)+(0.1,-0.1)$);
        \coordinate (right) at ($(left)+(1,0)$);
        \coordinate (mid) at ($(left)!0.5!(right)$);
        \coordinate (rlab) at ($(right)+(0.1,0)$);
        \coordinate (bleft) at ($(left)+(0,-8*\ysp)$);
        \coordinate (bright) at (rlab |- bleft);
        \coordinate (temp) at ($(origin)+(0,-3*\ysp)$);
        \draw[deme2] (left |- temp) -- (mid |- temp);
        \draw[deme1] (mid |- temp) -- (right |- temp);
        \coordinate (temp1) at ($(temp)+(0,-\ysp)$);
        \draw[deme1] (mid |- temp) -- (mid |- temp1) -- (right |- temp1);
        \coordinate (temp1) at ($(temp1)+(0,-\ysp)$);
        \draw[deme1] (mid |- temp) -- (mid |- temp1) -- (right |- temp1);
        \coordinate (temp1) at ($(temp1)+(0,-\ysp)$);
        \draw[deme1] (mid |- temp) -- (mid |- temp1) -- (right |- temp1);
        \node[text=darkgreen,font=\Huge] at (mid |- temp) {$\bullet$};
        \foreach \y in {2,7} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme1] (left |- temp) -- (right |- temp);
        }
        \foreach \y in {1,8} {
          \coordinate (temp) at ($(origin)+(0,-\y*\ysp)$);
          \draw[deme2] (left |- temp) -- (right |- temp);
        }
        \draw[timepoint] (mid) -- (mid |- bleft);
        \draw[axis,->] (bleft) -- (bright);
        \node[xshift=1ex] (tlab) at (bright.east) {$t$};
      \end{scope}
    \end{tikzpicture}
  }
\end{center}