### 2024-05-19 figs/seirs_events.tex
SEIRS table

\def\ysp{0.3}

\def\unbroken{%
  \resizebox{3em}{!}{%
    \begin{tikzpicture}[baseline=-0.5ex]
      \tikzstyle{deme1}=[color=darkblue, line width=2pt]
      \tikzstyle{deme1H}=[color=darkblue, line width=2pt, densely dotted]
      \tikzstyle{deme2}=[color=maize, line width=2pt]
      \tikzstyle{deme2H}=[color=maize, line width=2pt, densely dotted]
      \draw[deme1] (0,\ysp) -- (2,\ysp);
      \draw[deme2] (0,0) -- (2,0);
    \end{tikzpicture}%
  }%
}

\def\maizemaize{%
  \resizebox{3em}{!}{%
    \begin{tikzpicture}[baseline=-0.5ex]
      \tikzstyle{deme1}=[color=darkblue, line width=2pt]
      \tikzstyle{deme1H}=[color=darkblue, line width=2pt, densely dotted]
      \tikzstyle{deme2}=[color=maize, line width=2pt]
      \tikzstyle{deme2H}=[color=maize, line width=2pt, densely dotted]
      \tikzstyle{event}=[text=darkgreen, font=\Huge]
      \draw[deme2] (0,\ysp) -- (2,\ysp);
      \draw[deme1] (0,0) -- (2,0);
      \node[event] at (1,\ysp) {$\bullet$};
    \end{tikzpicture}%
  }%
}

\def\maizeblue{%
  \resizebox{3em}{!}{%
    \begin{tikzpicture}[baseline=-0.5ex]
      \tikzstyle{deme1}=[color=darkblue, line width=2pt]
      \tikzstyle{deme1H}=[color=darkblue, line width=2pt, densely dotted]
      \tikzstyle{deme2}=[color=maize, line width=2pt]
      \tikzstyle{deme2H}=[color=maize, line width=2pt, densely dotted]
      \tikzstyle{event}=[text=darkgreen, font=\Huge]
      \draw[deme2] (0,\ysp) -- (1,\ysp);
      \draw[deme1] (1,\ysp) -- (2,\ysp);
      \draw[deme1] (0,0) -- (2,0);
      \node[event] at (1,\ysp) {$\bullet$};
    \end{tikzpicture}%
  }%
}

\def\bluemaize{%
  \resizebox{3em}{!}{%
    \begin{tikzpicture}[baseline=-0.5ex]
      \tikzstyle{deme1}=[color=darkblue, line width=2pt]
      \tikzstyle{deme1H}=[color=darkblue, line width=2pt, densely dotted]
      \tikzstyle{deme2}=[color=maize, line width=2pt]
      \tikzstyle{deme2H}=[color=maize, line width=2pt, densely dotted]
      \tikzstyle{event}=[text=darkgreen, font=\Huge]
      \draw[deme1] (0,\ysp) -- (1,\ysp);
      \draw[deme2] (1,\ysp) -- (2,\ysp);
      \draw[deme2] (0,0) -- (2,0);
      \node[event] at (1,\ysp) {$\bullet$};
    \end{tikzpicture}%
  }%
}

\def\samplezero{%
  \resizebox{3em}{!}{%
    \begin{tikzpicture}[baseline=-0.5ex]
      \tikzstyle{deme1}=[color=darkblue, line width=2pt]
      \tikzstyle{deme1H}=[color=darkblue, line width=2pt, densely dotted]
      \tikzstyle{deme2}=[color=maize, line width=2pt]
      \tikzstyle{deme2H}=[color=maize, line width=2pt, densely dotted]
      \tikzstyle{event}=[text=darkgreen, font=\Huge]
      \tikzstyle{sample}=[text=royalblue, font=\Huge]
      \draw[deme1] (0,\ysp) -- (2,\ysp);
      \draw[deme2] (0,0) -- (1,0);
      \node[sample] at (1,0) {$\bullet$};
    \end{tikzpicture}%
  }%
}

\def\samplezeronix{%
  \resizebox{3em}{!}{%
    \begin{tikzpicture}[baseline=-0.5ex]
      \tikzstyle{deme1}=[color=darkblue, line width=2pt]
      \tikzstyle{deme1H}=[color=darkblue, line width=2pt, densely dotted]
      \tikzstyle{deme2}=[color=maize, line width=2pt]
      \tikzstyle{deme2H}=[color=maize, line width=2pt, densely dotted]
      \tikzstyle{event}=[text=darkgreen, font=\Huge]
      \tikzstyle{sample}=[text=royalblue, font=\Huge]
      \draw[deme1] (0,\ysp) -- (1,\ysp);
      \draw[deme2] (0,0) -- (2,0);
      \node[sample] at (1,\ysp) {$\bullet$};
    \end{tikzpicture}%
  }%
}

\def\sampleone{%
  \resizebox{3em}{!}{%
    \begin{tikzpicture}[baseline=-0.5ex]
      \tikzstyle{deme1}=[color=darkblue, line width=2pt]
      \tikzstyle{deme1H}=[color=darkblue, line width=2pt, densely dotted]
      \tikzstyle{deme2}=[color=maize, line width=2pt]
      \tikzstyle{deme2H}=[color=maize, line width=2pt, densely dotted]
      \tikzstyle{event}=[text=darkgreen, font=\Huge]
      \tikzstyle{sample}=[text=royalblue, font=\Huge]
      \draw[deme1] (0,\ysp) -- (2,\ysp);
      \draw[deme2] (0,0) -- (2,0);
      \node[sample] at (1,0) {$\bullet$};
    \end{tikzpicture}%
  }%
}

\def\sampleonenix{%
  \resizebox{3em}{!}{%
    \begin{tikzpicture}[baseline=-0.5ex]
      \tikzstyle{deme1}=[color=darkblue, line width=2pt]
      \tikzstyle{deme1H}=[color=darkblue, line width=2pt, densely dotted]
      \tikzstyle{deme2}=[color=maize, line width=2pt]
      \tikzstyle{deme2H}=[color=maize, line width=2pt, densely dotted]
      \tikzstyle{event}=[text=darkgreen, font=\Huge]
      \tikzstyle{sample}=[text=royalblue, font=\Huge]
      \draw[deme1] (0,\ysp) -- (2,\ysp);
      \draw[deme2] (0,0) -- (2,0);
      \node[sample] at (1,\ysp) {$\bullet$};
    \end{tikzpicture}%
  }%
}

\def\branchup{%
  \resizebox{3em}{!}{%
    \begin{tikzpicture}[baseline=-1ex]
      \tikzstyle{deme1}=[color=darkblue, line width=2pt]
      \tikzstyle{deme1H}=[color=darkblue, line width=2pt, densely dotted]
      \tikzstyle{deme2}=[color=maize, line width=2pt]
      \tikzstyle{deme2H}=[color=maize, line width=2pt, densely dotted]
      \tikzstyle{event}=[text=darkgreen, font=\Huge]
      \tikzstyle{sample}=[text=royalblue, font=\Huge]
      \draw[deme2] (0,0) -- (1,0);
      \draw[deme1] (1,0) -- (1,\ysp) -- (2,\ysp);
      \draw[deme2] (1,0) -- (1,-\ysp) -- (2,-\ysp);
      \node[event] at (1,0) {$\bullet$};
    \end{tikzpicture}%
  }%
}

\def\branchupnix{%
  \resizebox{3em}{!}{%
    \begin{tikzpicture}[baseline=-1ex]
      \tikzstyle{deme1}=[color=darkblue, line width=2pt]
      \tikzstyle{deme1H}=[color=darkblue, line width=2pt, densely dotted]
      \tikzstyle{deme2}=[color=maize, line width=2pt]
      \tikzstyle{deme2H}=[color=maize, line width=2pt, densely dotted]
      \tikzstyle{event}=[text=darkgreen, font=\Huge]
      \tikzstyle{sample}=[text=royalblue, font=\Huge]
      \draw[deme1] (0,0) -- (1,0);
      \draw[deme1] (1,0) -- (1,\ysp) -- (2,\ysp);
      \draw[deme2] (1,0) -- (1,-\ysp) -- (2,-\ysp);
      \node[event] at (1,0) {$\bullet$};
    \end{tikzpicture}%
  }%
}

\def\branchdown{%
  \resizebox{3em}{!}{%
    \begin{tikzpicture}[baseline=-1ex]
      \tikzstyle{deme1}=[color=darkblue, line width=2pt]
      \tikzstyle{deme1H}=[color=darkblue, line width=2pt, densely dotted]
      \tikzstyle{deme2}=[color=maize, line width=2pt]
      \tikzstyle{deme2H}=[color=maize, line width=2pt, densely dotted]
      \tikzstyle{event}=[text=darkgreen, font=\Huge]
      \tikzstyle{sample}=[text=royalblue, font=\Huge]
      \draw[deme2] (0,0) -- (1,0);
      \draw[deme2] (1,0) -- (1,\ysp) -- (2,\ysp);
      \draw[deme1] (1,0) -- (1,-\ysp) -- (2,-\ysp);
      \node[event] at (1,0) {$\bullet$};
    \end{tikzpicture}%
  }%
}

\def\branchdownnix{%
  \resizebox{3em}{!}{%
    \begin{tikzpicture}[baseline=-1ex]
      \tikzstyle{deme1}=[color=darkblue, line width=2pt]
      \tikzstyle{deme1H}=[color=darkblue, line width=2pt, densely dotted]
      \tikzstyle{deme2}=[color=maize, line width=2pt]
      \tikzstyle{deme2H}=[color=maize, line width=2pt, densely dotted]
      \tikzstyle{event}=[text=darkgreen, font=\Huge]
      \tikzstyle{sample}=[text=royalblue, font=\Huge]
      \draw[deme1] (0,0) -- (1,0);
      \draw[deme2] (1,0) -- (1,\ysp) -- (2,\ysp);
      \draw[deme1] (1,0) -- (1,-\ysp) -- (2,-\ysp);
      \node[event] at (1,0) {$\bullet$};
    \end{tikzpicture}%
  }%
}